# (PART\*) Bringing Your Own Data {-}


```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# Uploading from your local computer

In this example, we'll upload some genomic data into AnVIL.

TODO: Add information about what the data is

The starting point for bringing your own data to AnVIL is the Workspace Dashboard. At the bottom right, you'll find the full path to the Google Bucket information corresponding to your Workspace. You can click the clipboard icon on the right to copy the name of your Workspace Bucket. You will be able to see any uploaded files by clicking the "Open in browser" link.

```{r, echo=FALSE, fig.alt="Image shows a screenshot of the Workspace Dashboard. Google Bucket information, including the Google Bucket name, location, and 'Open in browser' link, at the bottom right of the screen is highlighted."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.gf5172664d7_0_142")
```

::: {.dictionary}
*Buckets** are the name of the containers used to store files and objects on Google Cloud. Everything you store on Google Cloud _must_ be in a bucket. Each bucket has its own unique name and location (URI). When we move data files into AnVIL workspaces, we use the URI to tell AnVIL where the data should be stored. (We can also use a URI to tell AnVIL where to find the data we want to upload.)

You can read more about Google Cloud buckets [here](https://docs.cloud.google.com/storage/docs/buckets)
:::


## Clone Workspace

## Identify bucket path

## Upload file 

## Check file status

## Bring file into a workspace

## Access Data Uploader

## Create a data collection

## Upload data collection

## Upload data table with metadata

## Summary

## Summary


# Uploading from the cloud

In this example, we'll upload some genomic data into AnVIL that is currently stored in the cloud (specifically, in a Google bucket).

::: {.dictionary}
*Buckets** are the name of the containers used to store files and objects on Google Cloud. Everything you store on Google Cloud _must_ be in a bucket. Each bucket has its own unique name and location (URI). When we move data files into AnVIL workspaces, we use the URI to tell AnVIL where the data should be stored. (We can also use a URI to tell AnVIL where to find the data we want to upload.)

You can read more about Google Cloud buckets [here](https://docs.cloud.google.com/storage/docs/buckets)
:::

We're going to upload some fastq files for a SARS-CoV-2 sample. The bucket we're accessing contains 5 samples: two compressed fastq files, a fasta file for a SARS-CoV-2 reference genome, and two uncompressed fastq files. The bucket ID (URI) is `fc-80d0e1cd-61e9-472f-b1bd-c6a8223bd1cd`. For this activity, you will retrieve the two uncompressed fastq files and upload them into your workspace.

```{r, echo=FALSE, fig.alt="Image shows the contents of a Google bucket used in the SARS-CoV-2 on Galaxy activity."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3afa908e330_0_178#slide=id.g3afa908e330_0_178")
```

## Step One: Create your workspace

The starting point for bringing your own data to AnVIL is the Workspace. Before you can do anything, you will need to create a workspace. Once you have logged into your AnVIL account, click on Workspaces in the left-side menu. You can open this menu by clicking the three line icon in the upper lefthand corner.

```{r, echo=FALSE, fig.alt="Once you have logged into your AnVIL account, click on Workspaces in the left-side menu. You can open this menu by clicking on the three line icon in the upper lefthand corner."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3afa908e330_0_182#slide=id.g3afa908e330_0_182")
```

Once you have opened the Workspace page, create a new workspace by clicking on the plus sign at the top.

```{r, echo=FALSE, fig.alt="Create a new workspace by clicking on the plus sign at the top."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3afa908e330_0_190#slide=id.g3afa908e330_0_190")
```

You should now see a pop-up window that lets you customize your new workspace. You will need to give your new workspace a unique name and assign it to a billing project. The "anvil-outreach" billing project is used here as an example, but you will not be able to assign it. You'll have to use one of your own billing projects. After filling out these two fields, click the "Quick Create Workspace" button to create a workspace without enabling sharing or additional security options.

```{r, echo=FALSE, fig.alt="You will need to give your new workspace a unique name and assign it to a billing project. The “anvil-outreach” billing project is used here as an example, but you won’t be able to use it. You’ll have to use one of your own. After filling out these two fields, click the “Quick Create Workspace” to create your workspace without enabling sharing or additional security options."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3afa908e330_0_197#slide=id.g3afa908e330_0_197")
```

You can read about Authorization Domains for workspace security in [this article](https://support.terra.bio/hc/en-us/articles/8527464803739-How-to-set-up-and-use-an-Authorization-Domain) at the AnVIL Portal.

Once you have created a workspace, AnVIL will take you to the Workspace Dashboard. 

## Step Two: Provision a Jupyter Notebook environment


## Step Three: Install gcloud storage

1.2. Run the following command using bash shells in your Terminal.

`curl https://sdk.cloud.google.com | bash`

1.3. Open a new bash shell in your terminal, or restart your shell using this command:

`exec -l $SHELL`

1.4.Authenticate by running the following:

`gcloud init`

::: {.notice}
Before downloading data using gcloud storage, verify your gcloud storage installation! Use the `ls` command to look at the buckets you can access.

Run `gcloud storage ls` to see all of the Cloud Storage buckets under the workspace project ID.

Run `gcloud storage ls -p [project name]` to list buckets for a specific project.
:::

Once gcloud storage is installed, you can perform command-line gcloud storage tasks to move/copy files between local storage and your workspace bucket (or any external bucket).

## Step Four: Run gcloud storage to copy/move files 

Now that you've installed gcloud storage, you can copy your data files into the Workspace. The command for this follows this pattern:

`gcloud storage cp <where_to_copy_data_from>/<file_name> <where_to_copy_data_to>  `
You will need the names of both the bucket where the data is currently stored _and_ the bucket the data is being moved to.

The data is currently stored at `fc-80d0e1cd-61e9-472f-b1bd-c6a8223bd1cd`.

The data is being moved to the Workspace bucket. The dashboard page of your new Workspace shows information about the Workspace on the right-hand side. At the bottom right, you'll find the full path to the Google Bucket information corresponding to your Workspace. This is the bucket where all information associated with this Workspace is stored.

You can click the clipboard icon on the right to copy the name of your Workspace Bucket.

```{r, echo=FALSE, fig.alt="The dashboard page of your new workspace shows information about the workspace on the right-hand side. You will want to copy the bucket name from here. This is the bucket where all information associated with this workspace is stored."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3afa908e330_0_205#slide=id.g3afa908e330_0_205")
```

For this example, the command to move the files is

`gcloud storage cp <where_to_copy_data_from>/<file_name> <where_to_copy_data_to>  `

## Step Five: Verify that files were moved to your workspace

On the Workspace Dashboard, you will be able to see any uploaded files by clicking the "Open in browser" link.

```{r, echo=FALSE, fig.alt="Image shows a screenshot of the Workspace Dashboard. Google Bucket information, including the Google Bucket name, location, and 'Open in browser' link, at the bottom right of the screen is highlighted."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.gf5172664d7_0_142")
```

You can also see any uploaded files by clicking the "Files" directory at the bottom left in the Data Tab.

```{r, echo=FALSE, fig.alt='Image shows a screenshot of the Workspace Data tab. The Files directory and link on the bottom left is highlighted.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.gf55fadc51c_0_3")
```


Data Tables provide a way to organize data and metadata, including URI links to storage buckets. These tables are a convenient way to organize input for analyses as well as tracking workflow outputs.

```{r, echo=FALSE, fig.alt="Image shows a schematic of the data storage locations in an AnVIL Workspace. The Data Table is highlighted with a number 'three'."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.gf982a3c0cd_0_0")
```



# Uploading from a remote cluster (HPC)

## Install `gsutil` on your local server

## Copy files

## Check file status

## Bring file into a workspace

## Summary


# Additional Resources

You can read documentation about bringing your own data to AnVIL on the [Portal](https://anvilproject.org/learn/find-data/bringing-your-own-data)

More details can be found in the [Terra documentation](https://support.terra.bio/hc/en-us/sections/360004147951)

## Information from Getting Started guide

## Browser: Upload Single Files

Click the "Files" directory at the bottom left of the Data Tab. Then click the "+" button in the bottom right corner of the screen. This will prompt a file browser on your local machine.

```{r, echo=FALSE, fig.alt='Image shows a screenshot of the Workspace Data tab. The plus button on the bottom right corner of the screen is highlighted.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.gf55fadc51c_0_12")
```

## Browser: Upload Folders

Click the ["Open in browser"](#bring-data-overview) link on the bottom right of the Workspace Dashboard Tab. This will open a new browser window or tab directed to your Workspace's Google Bucket on the Google Cloud Platform. 

```{r, echo=FALSE, fig.alt="Image shows a screenshot of the Workspace Dashboard. The 'Open in browser' link at the bottom right of the screen is highlighted."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.gf57004a098_0_1")
```

Here, you can upload files and manage your data and folders. You can also upload an entire folder by clicking on "UPLOAD FOLDER".

```{r, echo=FALSE, fig.alt="Image shows a screenshot of the Workspace Google Bucket on the Google Cloud Platform. The 'UPLOAD FOLDER' button is highlighted."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.gf57004a098_0_9")
```

## `gsutil`: Local to Cloud

`gsutil` is a Python application that lets you access Cloud Storage from the command line in a terminal. The terminal you use can be run on your local machine (local instance) or built into the Workspace Cloud Environment.

## Install `gsutil` on Your Local Computer or Local Server

[Cloud SDK](https://cloud.google.com/sdk/docs) is a set of tools that you can use to manage resources and applications hosted on Google Cloud. These tools include the `gsutil` command-line tool. 

1. Ensure you have a terminal available. 
    - MacOS and Linux users have a terminal application available by default. Terminal applications are also available through third party software, such as RStudio. 
    - Windows users should download a terminal application, such as [Putty](https://www.chiark.greenend.org.uk/~sgtatham/putty/latest.html).

1. Install Cloud SDK following the appropriate link below:
    - [Windows](https://cloud.google.com/sdk/docs/install#windows)
    - [MacOS](https://cloud.google.com/sdk/docs/install#mac)
    - [Linux](https://cloud.google.com/sdk/docs/install#linux)

1. Test that Cloud SDK has been successfully installed by typing `gsutil` in the terminal application prompt:

```
gsutil
```

If the installation was successful, you should see information about using `gsutil` that looks like the following:

```
Usage: gsutil [-D] [-DD] [-h header]... [-i service_account] [-m] [-o] [-q] [-u user_project] [command [opts...] args...]
```

If the installation was not successful, you should see a warning that `gsutil` was not found. Please return to the installation steps to ensure they have been completed correctly.

```
command not found: gsutil
```

## Copy Files From Your Local Computer to a Workspace Bucket

The `gsutil cp` command allows you to copy data from one machine to another. On your local machine's terminal, you should use the command in the following format:

```
gsutil cp where_to_copy_data_from/filename where_to_copy_data_to
```

**Example:** To copy the file `test.bam` located on your local computer at `users/name/data/` into the Workspace Bucket `gs://ab5-27x` on the cloud:

```
gsutil cp users/name/data/test.bam gs://ab5-27x
```

Remember that you can easily copy the Workspace Bucket ID using the clipboard button on the [Workspace Dashboard]({#bring-data-overview}). Please see the [`gsutil cp` documentation](https://cloud.google.com/storage/docs/gsutil/commands/cp) for more details, such as how to do parallel multi-threaded/multi-processing copying or copying an entire directory tree. The `gsutil cp` command can also be used to copy files from one Workspace Bucket to another (cloud-to-cloud copying).


