
```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```
# Uploading From Your Local Computer

TODO: ADD A SCENARIO WHICH WOULD REQUIRE PEOPLE TO UPLOAD THEIR DATA INTO ANVIL

In this example, we'll upload some genomic data that is stored on your local computer. You can use a file of your own, or download the files we are using in this example.

TODO: Add information about what the data is

::: {.notice}
_Genetics_

**Novice**: no genetics skills needed

_Programming skills_

**Novice**: no programming skills needed
:::

::: {.notice}
**What will this cost?**


:::

## Step One: Create your workspace

The starting point for bringing your own data to AnVIL is the workspace. Before you can do anything, you will need to create a workspace. Once you have logged into your AnVIL account, click on "Workspaces" in the left-side menu. You can open this menu by clicking the three line icon in the upper left hand corner.

```{r, echo=FALSE, fig.alt="Once you have logged into your AnVIL account, click on Workspaces in the left-side menu. You can open this menu by clicking on the three line icon in the upper lefthand corner."}
ottrpal::include_slide("")
```

Once you have opened the workspace page, create a new workspace by clicking on the plus sign at the top.

```{r, echo=FALSE, fig.alt="Create a new workspace by clicking on the plus sign at the top."}
ottrpal::include_slide("")
```

You should now see a pop-up window that lets you customize your new workspace. You will need to give your new workspace a unique name and assign it to a billing project. The "anvil-outreach" billing project is used here as an example, but you will not be able to assign it. You'll have to use one of your own billing projects. After filling out these two fields, click the "Quick Create Workspace" button to create a workspace without enabling sharing or additional security options.

```{r, echo=FALSE, fig.alt="You will need to give your new workspace a unique name and assign it to a billing project. The “anvil-outreach” billing project is used here as an example, but you don't have permission to use it. You’ll have to use one of your own. After filling out these two fields, click the “Quick Create Workspace” to create your workspace without enabling sharing or additional security options."}
ottrpal::include_slide("")
```

You can read about Authorization Domains for workspace security in [this article](https://support.terra.bio/hc/en-us/articles/8527464803739-How-to-set-up-and-use-an-Authorization-Domain) in the Terra documentation.

Once you have created a workspace, AnVIL will take you to the workspace dashboard. 

## Open the Data Uploader

When you switch to the Data tab in your Workspace, you'll see three headers on the left hand side: **TABLES**, **REFERENCE DATA**, and **WORKSPACE DATA**. At the moment all three of these sections are empty. We will use the Data Uploader tool to populate the **TABLES** section of the Workspace.

Click on the box in the upper left hand corner that says "Import Data". From the dropdown menu, you'll want to choose "Open data uploader".

TODO: Add image

When you first open the Data Uploader, you will be prompted to choose a workspace. Click on the workspace that you have just created.

TODO: Add image

## Upload data and metadata

After you have chosen your workspace, the Data Uploader then will then prompt you to create a data collection. Click on "Create a New Collection".

TODO: Add image

You will need to choose a name for your data collection. In our example, we chose to name the collection "covid_sequences". Make sure to pick something descriptive enough so that you'll know what the samples are later!

TODO: Add image

Next, click on the "Upload" button. The Data Uploader will prompt you to choose the files from your computer that you want to upload into your workspace bucket. Remember that every file must have a unique name!

TODO: Add image

After you upload your data files, you can always remove files. In this example, we accidentally uploaded a fasta file called "sars_membrane_protein_omicron.fasta". To remove it, check the box next to the file name and click the "Delete" button.

TODO: Add image

Once you have verified all the files that you want in your workspace bucket have been uploaded, click the "Next" button so that you can load the sample metadata. 

TODO: Add image

Data Uploader requires you to drag and drop your metadata file onto a bar at the bottom of the page.

TODO: Add image

Your metadata file should be saved as a tsv. The first line needs to be column names for the metadata. Additionally, the first column should be called "entity:sample_id". Each sample ID should match the name of the files you uploaded earlier. Remember to use tabs between each column.

TODO: Add image

The Data Uploader will display a table that displays your metadata. If it looks good, you can click "Create Table" to finish the file upload.

TODO: Add image

ADD A SECOND DATA COLLECTION INSTRUCTIONS

## Verify files were moved into your workspace

How do you know if your files were successfully transferred?

There are two ways to check! First, you should see information about the samples you loaded in the **TABLES** section of the Data tab.

```{r, echo=FALSE, fig.alt='TO ADD'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3c63398e684_33_33#slide=id.g3c63398e684_33_33")
```

You can also see any uploaded files by clicking the "Files" directory at the bottom left in the Data Tab.

```{r, echo=FALSE, fig.alt='Image shows a screenshot of the workspace Data tab. The Files directory and link on the bottom left is highlighted.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3c7f7fdd14a_0_24#slide=id.g3c7f7fdd14a_0_24")
```

```{r, echo=FALSE, fig.alt='Image shows a screenshot of the workspace Data tab. The Files directory and link on the bottom left is highlighted.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3c7f7fdd14a_0_27#slide=id.g3c7f7fdd14a_0_27")
```

If you click on the name of the file, you can also see the details of the file.

```{r, echo=FALSE, fig.alt='Image shows a screenshot of the details of the files we successfully transferred.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3c7f7fdd14a_0_30#slide=id.g3c7f7fdd14a_0_30")
```

This is a good time to check that the sizes of the files you transferred match the sizes of the original files.

## Summary

- Create a workspace
- Open Data Uploader
- Upload data and metadata (this requires creating a data collection)
- Verify that files have been transferred

## Additional Resources