# (PART\*) Finding Data on AnVIL {-}


```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

# AnVIL Data Library

The [Datasets Library](https://anvil.terra.bio/#library/datasets) is a good place to get started and familiarize yourself with existing data. Here, you can find curated datasets from thousands of participants. Some of these are open access (such as the [1000 Genomes dataset](https://anvil.terra.bio/#workspaces/anvil-datastorage/1000G-high-coverage-2019)) while others will require you to request access.

```{r, echo=FALSE, fig.alt='Image shows a screenshot of the Datasets Library landing page.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.gf5fa6f264a_0_15")
```

Taking a look at [Featured Workspaces](https://anvil.terra.bio/#library/showcase) can get you started quickly. Remember that when you [clone a Workspace](workspaces.html#clone-workspace), AnVIL automatically cross-links to the original data contained within the Data Tables. 

```{r, echo=FALSE, fig.alt='Image shows a screenshot of the Featured Workspaces landing page.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.gf5fa6f264a_0_24")
```

```{r, echo=FALSE, fig.alt='Image shows a screenshot of the Featured Workspaces tab on AnVIL. The featured tab is highlighted.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.gf5fa6f264a_0_7")
```

# AnVIL Dataset Catalog

The [AnVIL Dataset Catalog](https://anvilproject.org/data) displays key NHGRI datasets accessible in AnVIL, such as the CCDG (Centers for Common Disease Genomics), CMG (Centers for Mendelian Genomics), eMERGE (Electronic Medical Records and Genomics), as well as other relevant datasets. You will need to [coordinate access](https://anvilproject.org/learn/accessing-data/requesting-data-access) to controlled data.

```{r, echo=FALSE, fig.alt='Image shows a screenshot of the AnVIL Dataset Catalog website landing page.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.gf5fa6f264a_0_31")
```

# AnVIL Data Explorer

The [AnVIL Data Explorer](https://explore.anvilproject.org/datasets) enables faceted searches of open and managed access datasets hosted in AnVIL, making it easier for researchers to find and custom-build cohorts. 

```{r, echo=FALSE, fig.alt='Image shows a screenshot of the AnVIL Data Explorer website landing page.'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1H5onDH7cBLK2m7fCcJ6ZodAAQ3wtJO8tNc2rwptrTPM/edit#slide=id.g30d935bde8e_0_0")
```

# Additional Resources

More details can be found in the [Terra documentation](https://support.terra.bio/hc/en-us/articles/34594440468635-Finding-and-using-AnVIL-data).

Specific documentation about using the Data Explorer is [here](https://support.terra.bio/hc/en-us/articles/34596321862427-Part-2-Search-and-select-data-in-AnVIL-Data-Explorer)

Information on exporting data from the Data Explorer into an AnVIL workspace is [here](https://support.terra.bio/hc/en-us/articles/34607573660827-Part-3-Export-AnVIL-data-to-Terra-for-analysis)