
```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

#From Sequence Read Archive (SRA)

In this example, we'll bring some metagenomic data into AnVIL. This data comes from [this BioProject](https://www.ncbi.nlm.nih.gov/bioproject/PRJNA904247), which collected soil samples to study bacterial communities in tallgrass prairie. Bacteria play an important role in this ecosystem, but can be changed by disturbance, management, and the presence of herbivores.

We will bring this data into AnVIL from the **Sequence Read Archive**, or SRA. You can check out the [SRA website](https://www.ncbi.nlm.nih.gov/sra) to learn more:

> Sequence Read Archive (SRA) data, available through multiple cloud providers and NCBI servers, is the largest publicly available repository of high throughput sequencing data. The archive accepts data from all branches of life as well as metagenomic and environmental surveys. SRA stores raw sequencing data and alignment information to enhance reproducibility and facilitate new discoveries through data analysis. 

The SRA Data corresponding to this project is located [here](https://www.ncbi.nlm.nih.gov/Traces/study/?acc=SRP409181&o=acc_s%3Aa).

```{r, fig.align='center', echo = FALSE, fig.alt= "Microbiome diversity has many benefitial properties ranging soil and plant health.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1l0P0gFpsPkYG7blqJ_5JyYYlztJFZDD39CnIB4svrY8/edit#slide=id.g208b8f790dc_23_217")
```


::: {.notice}
_Genetics_

**Novice**: no genetics skills needed

_Programming skills_

**Intermediate**: no programming skills needed
:::

::: {.notice}
**What will this cost?**

You might hear new terms for moving data around in the cloud. **Ingress** is when data comes to you, similar to downloading a file or receiving an email with an attachment. **Egress** is sending the data to another resource, similar to uploading or sending an attached file via email. There is no fee for ingressing data to AnVIL from SRA.

There is also no cost for importing a workflow into your workspace, but there is a small cost for storing the data in an AnVIL bucket.
:::

## Step One: Create your workspace

The starting point for bringing your own data to AnVIL is the workspace. Before you can do anything, you will need to create a workspace. Once you have logged into your AnVIL account, click on "Workspaces" in the left-side menu. You can open this menu by clicking the three line icon in the upper left hand corner.

```{r, echo=FALSE, fig.alt="Once you have logged into your AnVIL account, click on Workspaces in the left-side menu. You can open this menu by clicking on the three line icon in the upper lefthand corner."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_36#slide=id.g3ce9c5d5921_0_36")
```

Once you have opened the workspace page, create a new workspace by clicking on the plus sign at the top.

```{r, echo=FALSE, fig.alt="Create a new workspace by clicking on the plus sign at the top."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_46#slide=id.g3ce9c5d5921_0_46")
```

You should now see a pop-up window that lets you customize your new workspace. You will need to give your new workspace a unique name and assign it to a billing project. The "anvil-outreach" billing project is used here as an example, but you will not be able to assign it. You'll have to use one of your own billing projects. After filling out these two fields, click the "Quick Create Workspace" button to create a workspace without enabling sharing or additional security options.

```{r, echo=FALSE, fig.alt="You will need to give your new workspace a unique name and assign it to a billing project. The “anvil-outreach” billing project is used here as an example, but you don't have permission to use it. You’ll have to use one of your own. After filling out these two fields, click the “Quick Create Workspace” to create your workspace without enabling sharing or additional security options."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_0#slide=id.g3ce9c5d5921_0_0")
```

You can read about Authorization Domains for workspace security in [this article](https://support.terra.bio/hc/en-us/articles/8527464803739-How-to-set-up-and-use-an-Authorization-Domain) in the Terra documentation.

Once you have created a workspace, AnVIL will take you to the workspace dashboard.

## Step Two: Import WDL workflow

In your new workspace, navigate to the Workflows tab. Any workflows that have been imported to the workspace will be displayed here. Since this is a new workspace, there are no workflows displayed. Click on "Find a Workflow" to get started.

```{r, fig.align='center', echo = FALSE, fig.alt= "After you navigate to the Workflows tab, click on the Find a Workflow box to get started.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_3#slide=id.g3ce9c5d5921_0_3")
```

A pop-up window will appear, asking if you'd like to import a workflow from either Dockstore.org or the Terra Workflow Repository. Click on the Dockstore.org box.

```{r, fig.align='center', echo = FALSE, fig.alt= "Choose the Dockstore.org option when asked from where you want to import a workflow.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_6#slide=id.g3ce9c5d5921_0_6")
```

A new browser window will open, showing a search page on Dockstore. Type "sra fetch" into the search bar and hit enter. By default the "WDL" option will be chosen as the workflow language. 

```{r, fig.align='center', echo = FALSE, fig.alt= "In the new browser window, use the search term 'sra fetch' to find a WDL.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_9#slide=id.g3ce9c5d5921_0_9")
```

In early March 2026, there were more than 400 WDLs deposited in Dockstore that helped with automating SRA workflows. For this vignette, we're looking for one called "pull_FASTQs_from_SRA_by_run" by user aofarrel. We have verified that this workflow still works as expected as of March 2026. Click on the WDL title.

```{r, fig.align='center', echo = FALSE, fig.alt= "You are looking for a WDL written by user aofarrel called 'pull_FASTQs_from_SRA_by_run' in Dockstore.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_12#slide=id.g3ce9c5d5921_0_12")
```

You should now see the page for the WDL you have chosen. WDLs from Dockstore can be imported into AnVIL with the click of a button. Choose the "AnVIL" box on the right side of the page.

```{r, fig.align='center', echo = FALSE, fig.alt= "Click on the AnVIL box on the righthand side of the page.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_15#slide=id.g3ce9c5d5921_0_15")
```

A new browser window for AnVIL will open and prompt you to choose the workspace you'd like to import the WDL to. Choose the name of the workspace that you created earlier. Notice that the workspace name starts with the billing project. In our example, the workspace is "anvil-outreach/import_data_sra". Your workspace will be differ and be named based on your billing project and the workspace name you chose. Once you have chosen the appropriate workspace, click Import.

```{r, fig.align='center', echo = FALSE, fig.alt= "Choose the name of the workspace that you created earlier. Notice that the workspace name starts with the billing project. In our example, the workspace is 'anvil-outreach/import_data_sra'.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_18#slide=id.g3ce9c5d5921_0_18")
```

After you have imported the workflow, you will see it as an option on the Workflow tab.

```{r, fig.align='center', echo = FALSE, fig.alt= "After you have imported the workflow, you will see it as an option on the Workflow tab.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_94#slide=id.g3ce9c5d5921_0_94")
```

## Step Three: Import sample TSV file

The WDL you have chosen requires a list of the SRA samples you want to import into your workspace. You can do this sample by sample, but a more efficient way is to save the sample IDs to a TSV file and let AnVIL run parallel workflows for each sample. 

Let's take a moment to get acquainted with the new file we'll be uploading.

First, download the samples file here. You might have to right-click and "Save as".

TODO: Add link to downloadable file

Next, open the file on your local machine. This is what it might look like in Microsoft Excel:

```{r, fig.align='center', echo = FALSE, fig.alt= "The samples we want to import from SRA are listed in rows in `samples.tsv, Make sure the column header is 'entity:sample_id'.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_24#slide=id.g3ce9c5d5921_0_24")
```

::: {.notice}
The column header `entity:sample_id` is important. `entity:` is required. `samples` becomes the name of the data table. So for example, if our header was `entity:reference_id`, a data table called "reference" would be created in AnVIL. If you didn't want to overwrite anything in the original "samples" table, you could change the column header. As long as none of the IDs are the same, no data will be overwritten. 
:::

Next, navigate to the Data tab in your workspace. Click on IMPORT DATA and select "Upload TSV".

```{r, fig.align='center', echo = FALSE, fig.alt= "The IMPORT DATA button and 'Upload TSV' option", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_27#slide=id.g3ce9c5d5921_0_27")
```

This opens a popup that looks like this:

```{r, fig.align='center', echo = FALSE, fig.alt= "The popup is titled Import Data Table and has the option to click to select a .tsv file", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_30#slide=id.g3ce9c5d5921_0_30")
```

Click to select your samples.tsv file. You will see a warning about potentially overwriting the existing entries. We know that none of the IDs in the new samples file overlap, so click START IMPORT JOB.

Now when we look at the TABLE in the Data tab, we can see that our samples have been added!

```{r, fig.align='center', echo = FALSE, fig.alt= "After we have imported the tsv file, the samples table in our workspace will be populated.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_33#slide=id.g3ce9c5d5921_0_33")
```

## Step Four: Launch workflow

We are finally ready to launch the workflow! Navigate to the Workflows tab and choose the WDL we imported previously.

```{r, fig.align='center', echo = FALSE, fig.alt= "On the Workflows tab, click on the WDL we imported earlier.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_71#slide=id.g3ce9c5d5921_0_71")
```

Most of the default options are fine for now, but we do want to tell the workflow to pull the sample names from the table we created earlier. Choose the radio button to select "Run workflow(s) with inputs defined by data table". You will be prompted to choose either "sample" or "sample_set" from the pull down menu. Either option will work, but for ease we will choose "sample". After making these changes, click the "Select Data" button under Step 2.

```{r, fig.align='center', echo = FALSE, fig.alt= "Choose the radio button to select "Run workflow(s) with inputs defined by data table". You will be prompted to choose either "sample" or "sample_set" from the pulldown menu. Either option will work, but for ease we will choose "sample". After making these changes, click the "Select Data" button under Step 2.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_68#slide=id.g3ce9c5d5921_0_68")
```

A new screen will open showing the sample IDs from the data table. Make sure the box next to each sample is checked. At the bottom of the screen, you will again have the option to import these samples as a sample set (this box will be checked by default). Uncheck this box, then choose the OK button in the lower right corner.

```{r, fig.align='center', echo = FALSE, fig.alt= "Make sure the samples you want to import have a checked box next to them, then choose OK.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_74#slide=id.g3ce9c5d5921_0_74")
```

After you have chosen the samples, the sample input field for the workflow will now read "this.sample_id". This just means that this workflow will run for each ID in the data table. The other input fields can be left as "optional". Click Save.

```{r, fig.align='center', echo = FALSE, fig.alt= "The sample input field for the workflow will be automatically populated after the samples have been chosen.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_77#slide=id.g3ce9c5d5921_0_774")
```

Once the inputs have been saved, you can start the workflow by clicking the Launch button.

```{r, fig.align='center', echo = FALSE, fig.alt= "Click the Launch button to start the workflow.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_80#slide=id.g3ce9c5d5921_0_80")
```

A pop-up window will be appear to verify that you want to launch the workflow. This window will also include details about any cost thresholds you did or did not set on your workflow. (We did not set any, but some workflows will incur charges. The box for setting a cost threshold is located just underneath the choose sample option on the workflow setup page.) Click Launch.

```{r, fig.align='center', echo = FALSE, fig.alt= "Click the Launch button to confirm that you want to start the workflow.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_83#slide=id.g3ce9c5d5921_0_83")
```

## Step Five: Check workflow progress

Details about all jobs, including workflow progress, can be accessed via the Submission History tab. You can click on a given job ID to access the specifics for that particular job.

```{r, fig.align='center', echo = FALSE, fig.alt= "Details about all submitted jobs can be found on the Submission History page.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_86#slide=id.g3ce9c5d5921_0_86")
```

At the bottom of the page, you can see the progress of each individual workflow submission. Because we had 10 sample IDs, we actually launched 10 individual workflows to retrieve the SRA data. These jobs will run in parallel. In early March 2026, we ran this particular workflow twice. The first time, most SRA samples were retrieved within 6 minutes, although the final two took 18 minutes. In our second trial, all samples were retrieved within 5 minutes.

For each individual run, you can click on the Job Manager icon for more details.

```{r, fig.align='center', echo = FALSE, fig.alt= "You can access details about individual runs by clicking on the job manager icon.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_89#slide=id.g3ce9c5d5921_0_89")
```

The job manager reports information such as job status (in this case, it succeeded), the input data, the output data, and a timing diagram of how long each step of the job took. If a job failed, it will also report the error that caused the failure.

```{r, fig.align='center', echo = FALSE, fig.alt= "The job manager reports information about job status, inputs, outputs, and errors.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_106#slide=id.g3ce9c5d5921_0_106")
```

## Step Six: Locate transferred data

Once all the jobs have completed successfully, you should check that you can find the transferred data. It's now located within the workspace bucket, which you can find by clicking on the File icon on the right side of the workspace Dashboard tab.

```{r, fig.align='center', echo = FALSE, fig.alt= "To access the transferred data, click on the File icon on the Dashboard tab.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_118#slide=id.g3ce9c5d5921_0_118")
```

The results of each individual run will be stored within a subfolder of the workflow submission name.

```{r, fig.align='center', echo = FALSE, fig.alt= "The results of each individual run will be stored within a subfolder of the workflow submission name.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_112#slide=id.g3ce9c5d5921_0_112")
```

You may have to click through several levels of directories to find the actual datafile. In this case, the results are within a folder labeled "shard-0".

```{r, fig.align='center', echo = FALSE, fig.alt= "You may have to click through several levels of directories to find the actual datafile. In this case, the results are within a folder labeled "shard-0".", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_115#slide=id.g3ce9c5d5921_0_115")
```

When you click on the "pull_results.txt" file, you have the option to download the FASTQ files the workflow retrieved from SRA. Downloading the files from AnVIL (also called "data egress") will cost a small amount. However, you do not have to download them in order to work with the data within an AnVIL computing environment.

```{r, fig.align='center', echo = FALSE, fig.alt= "When you click on the "pull_results.txt" file, you have the option to download the FASTQ files the workflow retrieved from SRA..", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_121#slide=id.g3ce9c5d5921_0_121")
```


## Summary

- Create a workspace
- Import a WDL workflow from Dockstore
- Upload a TSV containing sample IDs
- Launch the WDL
- Check the workflow progress 
- Locate the transferred data files


## Additional Resources


