
```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

#From Sequence Read Archive (SRA)

In this example, we'll bring some metagenomic data into AnVIL. This data comes from [this BioProject](https://www.ncbi.nlm.nih.gov/bioproject/PRJNA904247), which collected soil samples to study bacterial communities in tallgrass prairie. Bacteria play an important role in this ecosystem, but can be changed by disturbance, management, and the presence of herbivores.

We will bring this data into AnVIL from the **Sequence Read Archive**, or SRA. You can check out the [SRA website](https://www.ncbi.nlm.nih.gov/sra) to learn more:

> Sequence Read Archive (SRA) data, available through multiple cloud providers and NCBI servers, is the largest publicly available repository of high throughput sequencing data. The archive accepts data from all branches of life as well as metagenomic and environmental surveys. SRA stores raw sequencing data and alignment information to enhance reproducibility and facilitate new discoveries through data analysis. 

The SRA Data corresponding to this project is located [here](https://www.ncbi.nlm.nih.gov/Traces/study/?acc=SRP409181&o=acc_s%3Aa).

```{r, fig.align='center', echo = FALSE, fig.alt= "Microbiome diversity has many benefitial properties ranging soil and plant health.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1l0P0gFpsPkYG7blqJ_5JyYYlztJFZDD39CnIB4svrY8/edit#slide=id.g208b8f790dc_23_217")
```

::: {.dictionary}
You might hear new terms for moving data around in the cloud. **Ingress** is when data comes to you, similar to downloading a file or receiving an email with an attachment. **Egress** is sending the data to another resource, similar to uploading or sending an attached file via email. There is no fee for ingressing data to AnVIL from SRA.
:::


::: {.notice}
_Genetics_

**Novice**: no genetics skills needed

_Programming skills_

**Intermediate**: no programming skills needed
:::

::: {.notice}
**What will this cost?**


:::

## Step One: Create your workspace

The starting point for bringing your own data to AnVIL is the workspace. Before you can do anything, you will need to create a workspace. Once you have logged into your AnVIL account, click on "Workspaces" in the left-side menu. You can open this menu by clicking the three line icon in the upper left hand corner.

```{r, echo=FALSE, fig.alt="Once you have logged into your AnVIL account, click on Workspaces in the left-side menu. You can open this menu by clicking on the three line icon in the upper lefthand corner."}
ottrpal::include_slide("")
```

Once you have opened the workspace page, create a new workspace by clicking on the plus sign at the top.

```{r, echo=FALSE, fig.alt="Create a new workspace by clicking on the plus sign at the top."}
ottrpal::include_slide("")
```

You should now see a pop-up window that lets you customize your new workspace. You will need to give your new workspace a unique name and assign it to a billing project. The "anvil-outreach" billing project is used here as an example, but you will not be able to assign it. You'll have to use one of your own billing projects. After filling out these two fields, click the "Quick Create Workspace" button to create a workspace without enabling sharing or additional security options.

```{r, echo=FALSE, fig.alt="You will need to give your new workspace a unique name and assign it to a billing project. The “anvil-outreach” billing project is used here as an example, but you don't have permission to use it. You’ll have to use one of your own. After filling out these two fields, click the “Quick Create Workspace” to create your workspace without enabling sharing or additional security options."}
ottrpal::include_slide("")
```

You can read about Authorization Domains for workspace security in [this article](https://support.terra.bio/hc/en-us/articles/8527464803739-How-to-set-up-and-use-an-Authorization-Domain) in the Terra documentation.

Once you have created a workspace, AnVIL will take you to the workspace dashboard. 

## Step Two: Import WDL workflow

## Step Three: Import sample TSV file

## Step Four: Set up and launch workflow

## Step Five: Check workflow progress

## Step Six: Locate transferred data


How long did this take? In early March, did 10 samples. started at 1:20. Most completed within 6 minutes, but the last two workflows took 18 minutes (each sample fetch required a separate workflow, so we ran 10 workflows)

## Summary

- Create a workspace
- Import a WDL workflow from Dockstore
- Upload a TSV containing sample IDs
- Set up and launch the WDL
- Check the workflow progress 
- Locate the transferred data files


## Additional Resources


