
```{r, include = FALSE}
ottrpal::set_knitr_image_path()
```

#From Sequence Read Archive (SRA)

In this example, we'll bring some metagenomic data into AnVIL. This data comes from [this BioProject](https://www.ncbi.nlm.nih.gov/bioproject/PRJNA904247), which collected soil samples to study bacterial communities in tallgrass prairie. Bacteria play an important role in this ecosystem, but can be changed by disturbance, management, and the presence of herbivores.

We will bring this data into AnVIL from the **Sequence Read Archive**, or SRA. You can check out the [SRA website](https://www.ncbi.nlm.nih.gov/sra) to learn more:

> Sequence Read Archive (SRA) data, available through multiple cloud providers and NCBI servers, is the largest publicly available repository of high throughput sequencing data. The archive accepts data from all branches of life as well as metagenomic and environmental surveys. SRA stores raw sequencing data and alignment information to enhance reproducibility and facilitate new discoveries through data analysis. 

The SRA Data corresponding to this project is located [here](https://www.ncbi.nlm.nih.gov/Traces/study/?acc=SRP409181&o=acc_s%3Aa).

```{r, fig.align='center', echo = FALSE, fig.alt= "Microbiome diversity has many benefitial properties ranging soil and plant health.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1l0P0gFpsPkYG7blqJ_5JyYYlztJFZDD39CnIB4svrY8/edit#slide=id.g208b8f790dc_23_217")
```


::: {.notice}
_Genetics_

**Novice**: no genetics skills needed

_Programming skills_

**Intermediate**: no programming skills needed
:::

::: {.notice}
**What will this cost?**

You might hear new terms for moving data around in the cloud. **Ingress** is when data comes to you, similar to downloading a file or receiving an email with an attachment. **Egress** is sending the data to another resource, similar to uploading or sending an attached file via email. There is no fee for ingressing data to AnVIL from SRA.

There is also no cost for importing a workflow into your workspace, but there is a small cost for storing the data in an AnVIL bucket.
:::

## Step One: Create your workspace

The starting point for bringing your own data to AnVIL is the workspace. Before you can do anything, you will need to create a workspace. Once you have logged into your AnVIL account, click on "Workspaces" in the left-side menu. You can open this menu by clicking the three line icon in the upper left hand corner.

```{r, echo=FALSE, fig.alt="Once you have logged into your AnVIL account, click on Workspaces in the left-side menu. You can open this menu by clicking on the three line icon in the upper lefthand corner."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_36#slide=id.g3ce9c5d5921_0_36")
```

Once you have opened the workspace page, create a new workspace by clicking on the plus sign at the top.

```{r, echo=FALSE, fig.alt="Create a new workspace by clicking on the plus sign at the top."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_46#slide=id.g3ce9c5d5921_0_46")
```

You should now see a pop-up window that lets you customize your new workspace. You will need to give your new workspace a unique name and assign it to a billing project. The "anvil-outreach" billing project is used here as an example, but you will not be able to assign it. You'll have to use one of your own billing projects. After filling out these two fields, click the "Quick Create Workspace" button to create a workspace without enabling sharing or additional security options.

```{r, echo=FALSE, fig.alt="You will need to give your new workspace a unique name and assign it to a billing project. The “anvil-outreach” billing project is used here as an example, but you don't have permission to use it. You’ll have to use one of your own. After filling out these two fields, click the “Quick Create Workspace” to create your workspace without enabling sharing or additional security options."}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_0#slide=id.g3ce9c5d5921_0_0")
```

You can read about Authorization Domains for workspace security in [this article](https://support.terra.bio/hc/en-us/articles/8527464803739-How-to-set-up-and-use-an-Authorization-Domain) in the Terra documentation.

Once you have created a workspace, AnVIL will take you to the workspace dashboard.

## Step Two: Import WDL workflow

In your new workspace, navigate to the Workflows tab. Any workflows that have been imported to the workspace will be displayed here. Since this is a new workspace, there are no workflows displayed. Click on "Find a Workflow" to get started.

```{r, fig.align='center', echo = FALSE, fig.alt= "After you navigate to the Workflows tab, click on the Find a Workflow box to get started.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_3#slide=id.g3ce9c5d5921_0_3")
```

A pop-up window will appear, asking if you'd like to import a workflow from either Dockstore.org or the Terra Workflow Repository. Click on the Dockstore.org box.

```{r, fig.align='center', echo = FALSE, fig.alt= "Choose the Dockstore.org option when asked from where you want to import a workflow.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_6#slide=id.g3ce9c5d5921_0_6")
```

A new browser window will open, showing a search page on Dockstore. Type "sra fetch" into the search bar and hit enter. By default the "WDL" option will be chosen as the workflow language. 

```{r, fig.align='center', echo = FALSE, fig.alt= "In the new browser window, use the search term 'sra fetch' to find a WDL.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_9#slide=id.g3ce9c5d5921_0_9")
```

In early March 2026, there were more than 400 WDLs deposited in Dockstore that helped with automating SRA workflows. For this vignette, we're looking for one called "pull_FASTQs_from_SRA_by_run" by user aofarrel. We have verified that this workflow still works as expected as of March 2026. Click on the WDL title.

```{r, fig.align='center', echo = FALSE, fig.alt= "You are looking for a WDL written by user aofarrel called 'pull_FASTQs_from_SRA_by_run' in Dockstore.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_12#slide=id.g3ce9c5d5921_0_12")
```

You should now see the page for the WDL you have chosen. WDLs from Dockstore can be imported into AnVIL with the click of a button. Choose the "AnVIL" box on the righthand side of the page.

```{r, fig.align='center', echo = FALSE, fig.alt= "Click on the AnVIL box on the righthand side of the page.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_15#slide=id.g3ce9c5d5921_0_15")
```

A new browser window for AnVIL will open and prompt you to choose the workspace you'd like to import the WDL to. Choose the name of the workspace that you created earlier. Notice that the workspace name starts with the billing project. In our example, the workspace is "anvil-outreach/import_data_sra". Your workspace will be differ and be named based on your billing project and the workspace name you chose. Once you have chosen the appropriate workspace, click Import.

```{r, fig.align='center', echo = FALSE, fig.alt= "Choose the name of the workspace that you created earlier. Notice that the workspace name starts with the billing project. In our example, the workspace is 'anvil-outreach/import_data_sra'.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_18#slide=id.g3ce9c5d5921_0_18")
```

After you have imported the workflow, you will see it as an option on the Workflow tab.

```{r, fig.align='center', echo = FALSE, fig.alt= "After you have imported the workflow, you will see it as an option on the Workflow tab.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_94#slide=id.g3ce9c5d5921_0_94")
```

## Step Three: Import sample TSV file

The WDL you have chosen requires a list of the SRA samples you want to import into your workspace. You can do this sample by sample, but a more efficient way is to save the sample IDs to a TSV file and let AnVIL run parallel workflows for each sample. 

Let's take a moment to get acquainted with the new file we'll be uploading.

First, download the samples file here. You might have to right-click and "Save as".

TODO: Add link to downloadable file

Next, open the file on your local machine. This is what it might look like in Microsoft Excel:

```{r, fig.align='center', echo = FALSE, fig.alt= "The samples we want to import from SRA are listed in rows in `samples.tsv, Make sure the column header is 'entity:sample_id'.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_24#slide=id.g3ce9c5d5921_0_24")
```

::: {.notice}
The column header `entity:sample_id` is important. `entity:` is required. `samples` becomes the name of the data table. So for example, if our header was `entity:reference_id`, a data table called "reference" would be created in AnVIL. If you didn't want to overwrite anything in the original "samples" table, you could change the column header. As long as none of the IDs are the same, no data will be overwritten. 
:::

Next, navigate to the Data tab in your workspace. Click on IMPORT DATA and select "Upload TSV".

```{r, fig.align='center', echo = FALSE, fig.alt= "The IMPORT DATA button and 'Upload TSV' option", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_27#slide=id.g3ce9c5d5921_0_27")
```

This opens a popup that looks like this:

```{r, fig.align='center', echo = FALSE, fig.alt= "The popup is titled Import Data Table and has the option to click to select a .tsv file", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_30#slide=id.g3ce9c5d5921_0_30")
```

Click to select your samples.tsv file. You will see a warning about potentially overwriting the existing entries. We know that none of the IDs in the new samples file overlap, so click START IMPORT JOB.

Now when we look at the TABLE in the Data tab, we can see that our samples have been added!

```{r, fig.align='center', echo = FALSE, fig.alt= "After we have imported the tsv file, the samples table in our workspace will be populated.", out.width = '100%'}
ottrpal::include_slide("https://docs.google.com/presentation/d/1lSUfsg_oja-Iqq5pTFD1VOuR32GLUq2l-sCBYNO3OTg/edit?slide=id.g3ce9c5d5921_0_33#slide=id.g3ce9c5d5921_0_33")
```

## Step Four: Launch workflow

## Step Five: Check workflow progress

## Step Six: Locate transferred data


How long did this take? In early March, did 10 samples. started at 1:20. Most completed within 6 minutes, but the last two workflows took 18 minutes (each sample fetch required a separate workflow, so we ran 10 workflows)

## Summary

- Create a workspace
- Import a WDL workflow from Dockstore
- Upload a TSV containing sample IDs
- Launch the WDL
- Check the workflow progress 
- Locate the transferred data files


## Additional Resources


